Sustainable computational science:\\ the ReScience initiative

### 2017-07-13 draft.tex

\twocolumn[
  \maketitle
  \vspace{5mm}
  \noindent
  \textbf{Abstract.}
  %
  Computer science offers a large set of tools for prototyping, writing,
  running, testing, validating, sharing and replicating results, however
  computational science lags behind.  In the best case, authors may provide
  their source code as a compressed archive and they may feel confident their
  research is reproducible. But this is not exactly true. James Buckheit and
  David Donoho proposed more than 2 decades ago that an article about
  computational results is advertising, not scholarship. The actual scholarship
  is the full software environment, code, and data that produced the
  result. This implies new workflows, in particular in peer-reviews. Existing
  journals have been slow to adapt: source codes are rarely requested, hardly
  ever actually executed to check that they produce the results advertised in
  the article.
  %
  ReScience is a peer-reviewed journal that targets computational research and
  encourages the explicit replication of already published research, promoting
  new and open-source implementations in order to ensure that the original
  research can be replicated from its description. To achieve this goal, the
  whole publishing chain is radically different from other traditional
  scientific journals. ReScience resides on GitHub where each new
  implementation of a computational study is made available together with
  comments, explanations, and software tests.\\

@@

Confronted with this problem, a small but growing number of journals and
publishers have reacted by adopting explicit policies for data and software. Examples can be seen in the PLOS instructions on
\href{http://journals.plos.org/plosone/s/materials-and-software-sharing}{Materials
  and Software Sharing} and
on \href{http://journals.plos.org/plosone/s/data-availability}{Data Availability},
and in the
\href{https://elifesciences.org/elife-news/inside-elife-forking-software-used-elife-papers-github}{recent
  annoucement} by eLife on forking (creating a linked copy of) software used in eLife papers to GitHub.
Such policies help to ensure access to code and data in a well-defined format
\citep{perkel:2016} but this will not guarantee reproducibility nor
correctness.
At the educational and methodological levels, things have started to change with a
growing literature on best practices for making computations reproducible
\citep{sandve:2013, crook:2013, wilson:2014, halchenko:2015, janz:2015,
  hinsen:2015}.
  Related initiatives such as Software and Data Carpentry
\citep{wilson:2016} are of note since their goal is {\em to make
scientists more productive, and their work more reliable, by teaching them basic computing skills}.
Such best practices could be applied to already published research codebases as well, provided the original authors are willing to take on the challenge of 
re-implementing their software for the sake of better science.
Unfortunately, this is unlikely since the
incentives for doing such time-consuming work are low or nonexistent.
Furthermore, if the original authors made mistakes in their original
implementation, it seems likely that they will reproduce their mistakes
in any re-implementation.\\
% (just an educated guess, we have no data supporting this assertion)
%% At this point, the question is what do we do? Do we have to declare
%% the research to be lost once and for all because the associated software or
%% code is nowhere to be found / lost / damaged / malfunctioning / not runnable /
%% not usable / does not give original results / etc. We would doom ourselves by
%% doing so

@@

In theory, computation is a deterministic process and exact
reproduction should therefore be trivial. In reality, it is very
difficult to achieve because of the complexity of today's software
stacks and the tediousness of recording all interactions between a
scientist and a computer 
\citep[although a number of recent tools have attempted to automate such recording, e.g.][]{guo:2011, davison:2012, murta:2015}.
\citeauthor{Mesnard:2016} explain
\citep{Mesnard:2016} how difficult it can be to reproduce a two-year
old computation even though all possible precautions were taken at the
time to ensure reproducibility.  The most frequent obstacles are the
loss of parts of the software or input data, lack of a computing
environment that is sufficiently similar to the one used initially,
and insufficient instructions for making the software work. An
obstacle specific to numerical computations is the use of
floating-point arithmetic, whose rules are subject to slightly
different interpretations by different compilers and runtime support
systems. A large variety of research practices and support tools have
been developed recently to facilitate reproducible computations. For a
collection of recipes that have proven useful, see
\citet{kitzes:2017}.

@@

While some of the reasons for non-reproducibility are specific to each
scientific domain, our experience has shown that there are also some
common issues that can be identified.  Missing code and/or data,
undocumented dependencies, and inaccurate or imprecise description
appear to be characteristic of much non-reproducible work. Moreover,
these problems are not always easy to detect even for attentive
reviewers, as we discovered when some articles published in ReScience
turned out to be difficult to reproduce for someone else for exactly
the reasons listed above. ReScience reviewers are scientists working
in the same domain as the submitting authors, because familiarity with
the field is a condition for judging if a replication is
succesful. But this also means that our reviewers share a significant
common background with the authors, and that background often includes the
software packages and programming languages adopted by their
community.  In particular, if both authors and reviewers have
essential libraries of their community installed on their computers,
they may not notice that these libraries are actually dependencies of
the submitted code.  While solutions to this problem evidently exist
(ReScience could, for example, request that authors make their
software work on a standard computational environment supplied in the
form of a virtual machine), they represent an additional effort to
authors and therefore discourage them from submitting replication work to
ReScience.  Our goal is to develop a set of recommendations for
authors that represent a workable compromise between reproducibility
and implementation effort.\\

>>>>

\twocolumn[
  \maketitle
  \vspace{5mm}
  \noindent
  \textbf{Abstract.}
  %
  Computer science offers a large set of tools for prototyping, writing,
  running, testing, validating, sharing and replicating results, however
  computational science lags behind.  In the best case, authors may provide
  their source code as a compressed archive and they may feel confident their
  research is reproducible. But this is not exactly true. James Buckheit and
  David Donoho proposed more than two decades ago that an article about
  computational results is advertising, not scholarship. The actual scholarship
  is the full software environment, code, and data that produced the
  result. This implies new workflows, in particular in peer-reviews. Existing
  journals have been slow to adapt: source codes are rarely requested, hardly
  ever actually executed to check that they produce the results advertised in
  the article.
  %
  ReScience is a peer-reviewed journal that targets computational research and
  encourages the explicit replication of already published research, promoting
  new and open-source implementations in order to ensure that the original
  research can be replicated from its description. To achieve this goal, the
  whole publishing chain is radically different from other traditional
  scientific journals. ReScience resides on GitHub where each new
  implementation of a computational study is made available together with
  comments, explanations, and software tests.\\

@@

Confronted with this problem, a small but growing number of journals and
publishers have reacted by adopting explicit policies for data and software. Examples can be seen in the PLOS instructions on
\href{http://journals.plos.org/plosone/s/materials-and-software-sharing}{Materials
  and Software Sharing} and
on \href{http://journals.plos.org/plosone/s/data-availability}{Data Availability},
and in the
\href{https://elifesciences.org/elife-news/inside-elife-forking-software-used-elife-papers-github}{recent
  announcement} by eLife on forking (creating a linked copy of) software used in eLife papers to GitHub.
Such policies help to ensure access to code and data in a well-defined format
\citep{perkel:2016} but this will not guarantee reproducibility nor
correctness.
At the educational and methodological levels, things have started to change with a
growing literature on best practices for making computations reproducible
\citep{sandve:2013, crook:2013, wilson:2014, halchenko:2015, janz:2015,
  hinsen:2015}.
  Related initiatives such as Software and Data Carpentry
\citep{wilson:2016} are of note since their goal is {\em to make
scientists more productive, and their work more reliable, by teaching them basic computing skills}.
Such best practices could be applied to already published research codebases as well, provided the original authors are willing to take on the challenge of 
re-implementing their software for the sake of better science.
Unfortunately, this is unlikely since the
incentives for doing such time-consuming work are low or nonexistent.
Furthermore, if the original authors made mistakes in their original
implementation, it seems likely that they will reproduce their mistakes
in any re-implementation.\\
% (just an educated guess, we have no data supporting this assertion)
%% At this point, the question is what do we do? Do we have to declare
%% the research to be lost once and for all because the associated software or
%% code is nowhere to be found / lost / damaged / malfunctioning / not runnable /
%% not usable / does not give original results / etc. We would doom ourselves by
%% doing so

@@

In theory, computation is a deterministic process and exact
reproduction should therefore be trivial. In reality, it is very
difficult to achieve because of the complexity of today's software
stacks and the tediousness of recording all interactions between a
scientist and a computer 
\citep[although a number of recent tools have attempted to automate such recording, e.g.][]{guo:2011, davison:2012, murta:2015}.
\citeauthor{Mesnard:2016} explain
\citep{Mesnard:2016} how difficult it can be to reproduce a two-year-old
computation even though all possible precautions were taken at the
time to ensure reproducibility.  The most frequent obstacles are the
loss of parts of the software or input data, lack of a computing
environment that is sufficiently similar to the one used initially,
and insufficient instructions for making the software work. An
obstacle specific to numerical computations is the use of
floating-point arithmetic, whose rules are subject to slightly
different interpretations by different compilers and runtime support
systems. A large variety of research practices and support tools have
been developed recently to facilitate reproducible computations. For a
collection of recipes that have proven useful, see
\citet{kitzes:2017}.

@@

While some of the reasons for non-reproducibility are specific to each
scientific domain, our experience has shown that there are also some
common issues that can be identified.  Missing code and/or data,
undocumented dependencies, and inaccurate or imprecise description
appear to be characteristic of much non-reproducible work. Moreover,
these problems are not always easy to detect even for attentive
reviewers, as we discovered when some articles published in ReScience
turned out to be difficult to reproduce for someone else for exactly
the reasons listed above. ReScience reviewers are scientists working
in the same domain as the submitting authors, because familiarity with
the field is a condition for judging if a replication is
successful. But this also means that our reviewers share a significant
common background with the authors, and that background often includes the
software packages and programming languages adopted by their
community.  In particular, if both authors and reviewers have
essential libraries of their community installed on their computers,
they may not notice that these libraries are actually dependencies of
the submitted code.  While solutions to this problem evidently exist
(ReScience could, for example, request that authors make their
software work on a standard computational environment supplied in the
form of a virtual machine), they represent an additional effort to
authors and therefore discourage them from submitting replication work to
ReScience.  Our goal is to develop a set of recommendations for
authors that represent a workable compromise between reproducibility
and implementation effort.\\
